---
title       : Introduction to R
subtitle    : with unique features
author      : Tong He
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [nvd3,morris]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Outline

- R: Born for statistics.
- Thousands of packages for data analysis
- Reproducible report
    - knitr
    - slidify
    - rmarkdown
- Data Visualization
    - ggplot2
    - rCharts
- Web application
    - shiny

--- .segue .dark

## The GNU-R project

---

## R: Born for statistics

History:

- From the S language, which is inspired by Scheme.
- Created by [Ross Ihaka](http://en.wikipedia.org/wiki/Ross_Ihaka) and [Robert Gentleman](http://en.wikipedia.org/wiki/Robert_Gentleman_%28statistician%29).

---

## R: Born for statistics

R is strong for statistical missions. 

- Strong matrix calculation ability
- Native support for statistics
- Abundant third-party packages

---

## R: Born for statistics

```{r}
groupA = rnorm(100,5,3)
groupB = rnorm(25,10,3)
groupA
groupB
```

---

## R: Born for statistics

```{r}
t.test(groupA,groupB,alternative = c("less"))
```

---

## R: Born for statistics

```{r}
x = rnorm(100)
y = 5*x+rnorm(100,0,1)
lm.model = lm(y~x)
lm.model
```

---

## R: Born for statistics

```{r}
summary(lm.model)
```

---

## R: Born for statisitics

```{r,fig.align='center',fig.height=6,fig.width=8}
plot(x,y,pch=20)
abline(lm.model,col=2)
```

---

## Third party packages

- [CRAN](http://www.r-project.org/)(The Comprehensive R Archive Network) is managing over 5,000 R packages.
    - Strict quality control
    - [Task Views](http://cran.r-project.org/web/views/)
    - My package [xgboost](http://cran.r-project.org/web/packages/xgboost/index.html)
- [R-Forge](https://r-forge.r-project.org/) is an SVN repository for packages.
- [Bioconductor](http://www.bioconductor.org/) focus on biostat and bioinfo packages.
- [Github](http://github.com) more flexible.

---

## Third party packages

To install a package, say `xgboost`, on CRAN, just type 

```{r, eval=FALSE}
install.packages('xgboost')
```

To install `xgboost` from github, we need 

```{r,eval=FALSE}
require(devtools)
install_github('xgboost','tqchen',subdir='R-package')
```

--- &twocol

## The community and environment

*** =left

The community

- Mailing lists
- Stackoverflow / Cross Validation
- R conference

*** =right

The environment

- IDE: RStudio
  - Server version for collaboration
  - Many fancy stuffs later
- Commercial version
  - Revolution

--- .segue .dark

## Unique features in R

---

## Vectorization

```{r}
a = 1:1000000
b = 1:1000000
system.time({a+b})
system.time({
  for (i in 1:length(a))
    a[i]+b[i]
})
```

Let C to do loop.

---

## Vectorization

R is also strong in (sparse) matrix computation.

```{r}
library('Matrix')
 
m1 = matrix(0, nrow = 1000, ncol = 1000)
m2 = Matrix(0, nrow = 1000, ncol = 1000, sparse = TRUE)
 
object.size(m1)
object.size(m2)
```

---

## Vectorization

```{r}
m1[500, 500] <- 1
m2[500, 500] <- 1
 
object.size(m1)
object.size(m2)
```

---

## Vectorization

```{r, eval=FALSE}
m2 %*% rnorm(1000)
m2 + m2
m2 - m2
t(m2)
```

---

## Vectorization

R can also take advantage from the parallelization of *BLAS libraries. There are several choices:

- OpenBLAS
- ATLAS
- HiPLAR

[Here](http://www.stat.cmu.edu/~nmv/2013/07/09/for-faster-r-use-openblas-instead-better-than-atlas-trivial-to-switch-to-on-ubuntu/)'s a tutorial for apt-get style installation on Ubuntu.

---

## Reproducible Report

Insert code and result in a slide is painful.

- Copy & paste
- Formatting
- Update code and report seperately

---

## Reproducible Report

```{r,fig.align='center',fig.width=8,fig.height=5}
rnums = rnorm(10000)
head(rnums)
hist(rnums,breaks=50)
```

---

## Reproducible Report

This is supported by the package [`knitr`](http://yihui.name/knitr/):

- Markdown-based grammar with HTML codes as extension
- Webpage based output

We'll see the 'Webpage based output' is important.

--- &twocol

## Reproducible Report

This slide is produced by the package [`slidify`](https://github.com/ramnathv/slidify).

- Based on `knitr`
- Abundant templates
- Easy to share

---

## Reproducible Report

Now I write all the assignments with `rmarkdown`:

- Based on `knitr`
- Allowing formatting and templates
- Using `pandoc` as the backend
    - PDF
    - HTML
    - Word(yes, even word)

--- &twocol

## Data Visualization

Data visualization is sometimes important. Compare these two graphs:

*** =left

```{r,echo=FALSE,fig.height=5, fig.width=6, message=FALSE}
set.seed(6298)
require(ggplot2)
diamonds_small <- diamonds[sample(nrow(diamonds), 1000), ]
hist(diamonds_small$price,breaks=50)
```

*** =right

```{r,echo=FALSE,fig.height=5, fig.width=6, message=FALSE}
hist_cut <- ggplot(diamonds_small, aes(x=price, fill=cut))
hist_cut + geom_bar()
```

--- &twocol

## Data Visualization

Or we can have other variations:

*** =left

```{r,echo=FALSE,fig.height=5, fig.width=6, message=FALSE}
hist_cut + geom_bar(position="fill")
```

*** =right

```{r,echo=FALSE,fig.height=5, fig.width=6, message=FALSE}
hist_cut + geom_density(alpha = 0.2)
```

---

## Data Visualization

Sometimes static pictures are not enough:

```{r,echo=FALSE, message=FALSE, comment=NA}
require(rCharts)
require(knitr)
options(RCHART_WIDTH = 800, RCHART_HEIGHT = 500, rcharts.cdn=FALSE)
```

```{r,echo=FALSE,results = 'asis'}
hair_eye_male <- subset(as.data.frame(HairEyeColor), Sex == "Male")
n1 <- nPlot(Freq ~ Hair, group = "Eye", data = hair_eye_male, 
  type = 'multiBarChart')
n1$show('inline',include_assets = TRUE,cdn=FALSE)
```

---

## Data Visualization

```{r,echo=FALSE,results='asis'}
data(economics, package = "ggplot2")
econ <- transform(economics, date = as.character(date))
m1 <- mPlot(x = "date", y = c("psavert", "uempmed"), type = "Line", data = econ)
m1$set(pointSize = 0, lineWidth = 1)
m1
```

--- .segue .dark

## Web Application

---

## Web Application

Integrate all the previous features, we can merge them into a web app.

Yes, we can build a web app in R!

---

## Web Application

Examples: 

1. [Kmeans on iris data](http://shiny.rstudio.com/gallery/kmeans-example.html)
2. [Combining rCharts](http://shiny.rstudio.com/gallery/nvd3-line-chart-output.html)
3. [Wordcloud](http://shiny.rstudio.com/gallery/word-cloud.html)

---

## Web Application

Advanced Examples:

1. [imgSVD](https://hetong.shinyapps.io/imgsvd)
2. [Chat Room](http://shiny.rstudio.com/gallery/chat-room.html)
3. [Downloadable report genereation](http://shiny.rstudio.com/gallery/download-knitr-reports.html)
4. [Dynamic Clustering](http://shiny.rstudio.com/gallery/dynamic-clustering.html)
5. [Multi-language-character encoding](http://shiny.rstudio.com/gallery/unicode-characters.html)
6. [Personalization and authencation](http://shiny.rstudio.com/gallery/personalized-ui.html)

Enjoy more from the [gallery](http://shiny.rstudio.com/gallery/).

---

## Summarize

One of the important feature is the 'Webpage based output'. 

Tools as 

- ggplot2/rCharts
- knitr/slidify/rmarkdown
- shiny

enable us directly share R code and trustful result around the world.

--- &twocol

## Why does R need them

Why these are features unique?

*** =left

Academical application:

- Teaching
- Research

Demonstrate the numbers and graphs from the computation is important. And the need for sharing and publishing a report/lecture/tutorial is increasing.

*** =right

Industrial application:

- Internal communication
- Business report
- Automatic report generation

R is mostly used with data related missions. Usually clients would like to get information from slides with beautiful pictures. The report system connects the demonstration and data manipulation.

--- .segue .dark

## FAQ











































